---
title: "Psychophysical model recovery using Bayesian brms"
format: html
---

TO-DO:

  - Make it multilelvel, with multiple participants

To get started, we load the required packages.

```{r}
#| warning: false 
#| output: false
rm(list = ls())

library(tidyverse)
library(brms)

source("../R/simulate_data.R") #Load my needed custom function
source("../R/psychometric_function.R") #Load my needed custom function

set.seed(999) #ensures reproducibility for testing

break_brms_by_making_numTargets_numeric<- FALSE #Change factor to numeric, right before doing formula fit, so 
   #I know that it's brms that's the problem rather than the rest of my code
```

# Create simulated trials

Set up simulated experiment design.

```{r}
#| echo: true 
numTargetsConds<- c("two","three")
numSubjects<- 25
trialsPerCondition<- 8
laboratories<- c("Holcombe","Roudaia")
#Array of speeds (not very realistic because mostly controlled by a staircase in actual experiment)
speeds<-seq(.02,1.8, length.out = 12) # trials at 12 different speeds between .02 and 1.8

```

In order to build and test our model in brms, we must first create a simulated data set that is similar to our actual experiment data. This allows us to confirm the brms model is working and successfully recovers the parameters we set before applying it to our real experimental data that has unknown parameter values. In the actual data, there will be many group-wise differences in location and scale parameters. The following simulated data only has explicit differences between the $\eta$ (location) of the two age groups (older vs younger).


```{r}
#| echo: false 

trials <- generate_conditions(laboratories,numTargetsConds,numSubjects,trialsPerCondition,speeds)


#Print number of unique values of each column
#print('Number of values for each factor:')
numValsPerFactor<- trials |> summarise(across(everything(), ~ n_distinct(.))) |>
                              pivot_longer(everything())
print( numValsPerFactor )

```
Create between-subject variability within each group, to create a potential multilevel modeling advantage. Can do this by passing the participant number to the location-parameter calculating function. Because that function will be called over and over, separately, the location parameter needs to be a deterministic function of the participant number. So it needs to calculate a hash or something to determine the location parameter. E.g. it could calculate the remainder, but then the location parameters wouldn't be centered on the intended value for that condition. To do that, I think I need to know the number of participants in the condition, n, and number them 1..n. Then I can give e.g. participant 1 the extreme value on one side of the condition-determined value and give participant n the extreme value on the other side.
So I number participants separately even within age*gender to maintain the age and gender penalties, but not within speed, of course. Also not within targetLoad or obj_per_ring.

Because the present purpose of numbering participants is to inject the right amount of between-participant variability, I will number the participants with a range of numbers that has  unit standard deviation. Will call this column "subjStandardized".

```{r}
#| echo: false

  #Renumber subjWithinCond to be standardized, centered on zero and have unit standard deviation, so that that person's unique params 
  # can be assigned by multiplying subjWithinCond by the variance, because then the standard
  # deviation 
  # adding it to the designated mean val.
  center_on_zero_and_standardize_std <- function(x,vals) {
    centered <- (x - mean(vals)) / sd(vals)
    return(centered)
  }

  trials<-trials |> 
    mutate( subj_standardzd = center_on_zero_and_standardize_std(subjWithinGroup,
                                                          unique(trials$subjWithinGroup))
          )
  
  #In case I need trial number, so far I only have a trialThisCond column
  #Calculate a trial number numbering the entirety of the trials the subject is given
  #Assume the within-participant factors are obj_per_ring,targetLoad, and speed
  trials <- trials |> group_by(lab,numTargets,age_group,gender,subjWithinGroup) |> 
                      mutate(trial = row_number())
```

Choose values for psychometric function for younger and older

```{r}
#| echo: true 
lapse <- 0.05
sigma <- 0.2

location_param_base<-1.7 #location_param_young_123targets <- c(1.7,1.0,0.8)
target_penalty<-0.15 #penalty for additional target
age_penalty <- 0.2#0.4 #Old people have worse limit by this much
gender_penalty <- 0.09#0.1 #Female worse by this much
Holcombe_lab_penalty <- 0.07#0.2
#Set parameters for differences between Ss in a group
eta_between_subject_sd <- 0.2 

#Using above parameters, need function to calculate a participant's location parameter
#Include optional between_subject_variance
location_param_calculate<- function(numTargets,age_group,gender,lab,
                                    subjStandardized,eta_between_subject_sd) {
  
  base_location_param <- location_param_base#location_param_young_123targets[targetLoad]
  #calculate offset for this participant based on desired sigma_between_participant_variance
  location_param <- base_location_param + 
                    subjStandardized * eta_between_subject_sd
  
  after_penalties <-location_param - 
                      if_else(lab=="Holcombe",1,0) * Holcombe_lab_penalty -
                      if_else(age_group=="older",1,0) * age_penalty -
                      if_else(numTargets=="three",1,0) * target_penalty -
                      if_else(gender=="F",1,0) * gender_penalty
  return (after_penalties)
}

#Need version with fixed eta_between_subject_sd for use in mutate
location_param_calc_for_mutate<- function(numTargets,age_group,gender,lab,
                                               subjStandardized) {
  location_param_calculate(numTargets,age_group,gender,lab,
                           subjStandardized, eta_between_subject_sd)
}
  
```


Using the psychometric function, simulate whether participant is correct on each trial or not, and add that to the simulated data.

```{r}
#| echo: false

data_simulated<- trials

#Add column for the probability of each trial being correct based on our psychometric function
#then use that to add column to generate if the subject got the trial correct
data_simulated <- data_simulated %>%
  mutate(
    chance_rate = 1/obj_per_ring,
    location_param = location_param_calc_for_mutate(numTargets,age_group,gender,
                                                    lab,subj_standardzd),
    p_correct = psychometric_function(1/obj_per_ring,lapse,speed,location_param,sigma),
    correct = rbinom(n=length(p_correct), size=1, prob=p_correct)
  )

#Create qualitative target load in case it being numeric is the problem for brms
#data_simulated<- data_simulated |> mutate( target_load = 
#                                      if_else(targetLoad==2,"less","more") )

```    
    
# Plot data

upper_bound = 1 - L*(1-C)

```{r}
#| echo: false

#Calculate threshold (eta) for each group*condition

#data_one_subject_group <- data_simulated |> 
#  filter(age_group=="younger",gender=="M",lab=="Holcombe")
#data_one_condition <- data_one_subject_group |> filter(targetLoad==2)

gg<- ggplot(data_simulated, #data_one_condition, 
            aes(x=speed,y=p_correct,linetype=age_group,color=factor(numTargets))) +
  #stat_summary(fun=mean,geom="point") +
  stat_summary(fun=mean,geom="line")  +
  facet_grid(lab~obj_per_ring) +
  labs(x = "Speed (revolutions per second)",
        y = "P(Correct)",
        title = "Simulated data") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank()) #remove gridlines

#Also show between-subject variability
range <- function(x) {
  data.frame(
    y = mean(x), ymin = min(x), ymax = max(x)
  )
}
SDrange <- function(x) {
  data.frame(
    y = mean(x),
    ymin = mean(x) - 0.5*sd(x),
    ymax = mean(x) + 0.5*sd(x)
  )
}

#Can show variability as range or as CI
gg<-gg+stat_summary( aes(group=interaction(age_group,numTargets)),
                     fun.data=range, geom="ribbon", color=NA, fill="grey80", alpha=0.2 )
#Add confidence interval ribbon to show variability.
#gg<-gg+stat_summary( aes(group=interaction(age_group,targetLoad)),
#                    fun.data = mean_cl_normal, fun.args = list(conf.int = 0.95),
#                    geom="ribbon", fill="grey80", color=NA, alpha = 0.5)
#Show floor  and ceiling with lines
gg<- gg + geom_hline( aes(yintercept = chance_rate),
                      colour = "purple", alpha=0.2 ) +
     geom_hline( aes(yintercept = 1-lapse*(1-chance_rate)), 
                 colour = "yellow3", alpha=0.8 )

#Add threshold (location_parameter) lines by calculating average for each group of subjects,
#including midpoint (threshold percent correct level)
calc_midpoint <- function( chance_rate, lapse ) {
  (chance_rate + 1-lapse*(1-chance_rate)) / 2
}

#calculate the threshold percent correct level, which differs by num_objects as well as lapse
loc_param_each_group <- data_simulated |>
        group_by(age_group, lab, numTargets) %>%
        summarise(location_param = mean(location_param, na.rm = TRUE),
                  #Calculate midpoint, but I'm actually not sure if f(location_param) = midpoint
                  midpoint =  mean( calc_midpoint(chance_rate,lapse) ), 
                  .groups = "drop")

gg<-  gg+ geom_segment(data = loc_param_each_group,
                   aes(x = location_param, xend = location_param,
                       y = 0, yend = midpoint,
                       color=factor(numTargets)),
                       linetype = "dashed",alpha=.4)#, color = "grey3")
#This removes the default padding/margin below the y-axis, so the plot area starts exactly at y = 0.
gg<- gg + scale_y_continuous(limits=c(0,1),expand = c(0, 0))
show(gg)


#geom_hline(aes(yintercept = 1-lapse*(1-chance_rate), colour = "Upper Bound"), linetype = "dashed") +
    # geom_hline(aes(yintercept = 0.25 , colour = "Lower Bound"), linetype = "dashed") +
    # geom_vline(aes(xintercept = 0.9, colour = "eta")) +
    # geom_line(aes(x = speed, y = probability_correct,
    #               colour = "Probability Correct")) +
    # theme_light() +
    # lims(x = c(0,2.5), y = c(0,1)) +
    # scale_colour_manual(values = c("Upper Bound" = "blue", "Lower Bound" = "red", "eta" = "yellow", "Probability Correct" = "black")) +

```  

# Setting up our Model in brms

Setting a model formula in brms allows the use of multilevel models, where there is a hierarchical structure in the data. But at this point we haven't made the model multi-level as we have been concentrating on the basics of brms.

The bf() function of brms allows the specification of a formula. The parameter can be defined by population effects, where the parameter's effect is fixed, or group level effects where the parameter varies with a variable such as age. The "family" argument is a description of the response distribution and link function that the model uses. For more detailed information on setting up a formula and the different arguments in BRMS see<https://paulbuerkner.com/brms/reference/brmsformula.html>

The model we used is based off our psychometric function used to generate the data mentioned previously. The only explicitly-coded difference in our simulated data is in the location parameter of older vs younger. Thus, in addition to the psychometric function, we allowed $\eta$ and $\log(\sigma)$ to vary by age group in the model. Because the psychometric function doesn't map onto a canonical link function, we use the non-linear estimation capability of brms rather than linear regression with a link function.

*Alex's note: Using the nonlinear option is also what allowed us to set a prior on the thresholds  $\eta$, because we could then parametrize the function in terms of the x-intercept, whereas with the link-function approach, we are stuck with the conventional parameterization of a line, which [has a term for the y-intercept but not the x-intercept](https://bsky.app/profile/did:plc:kynaetyuzsp46xejc6mzpjle/post/3lg5lpartzs2z) *


```{r}
my_formula <- brms::bf( 
   correct ~ chance_rate + (1-chance_rate - lapseRate*(1-chance_rate)) * Phi(-(speed-eta)/exp(logSigma))
  ) 
my_formula <- my_formula$formula
  
my_brms_formula <- brms::bf(
  correct ~ chance_rate + (1-chance_rate - lapseRate * (1-chance_rate))*Phi(-(speed-eta)/exp(logSigma)), 
  eta ~ lab + age_group + numTargets + gender,
  lapseRate ~ 1, #~1 estimates intercept only
  logSigma ~ 1,#age_group,
  family = bernoulli(link="identity"), #Otherwise the default link 'logit' would be applied
  nl = TRUE #non-linear model
)
```

# Set priors

See [visualize_and_select_priors.html](visualize_and_select_priors.html) for some motivation and visualisation.

```{r}
#| echo: true 

my_priors <- c(
  brms::set_prior("beta(2,33.33)", class = "b", nlpar = "lapseRate", lb = 0, ub = 1),
  brms::set_prior("uniform(0, 2.5)", class = "b", nlpar = "eta", lb = 0, ub = 2.5),
  brms::set_prior("uniform(-3, 1.6)", class = "b", nlpar = "logSigma", lb = -2, ub = 1.6) 
)

```

# Set priors

See [visualize_and_select_priors.html](visualize_and_select_priors.html) for motivation.

```{r}
#| echo: false 
lapse_param1<- 2
lapse_param2<- 33.33

eta_param1<- 0
eta_param2<- 2.5

logsigma_param1<- -3
logsigma_param2<- 1.6

#brms can't evaluate parameters in the prior setting so one has to resort to sprintf-ing a string
lapseRate_prior_distribution<- sprintf("beta(%s, %s)", lapse_param1, lapse_param2)
eta_prior_distribution<-  sprintf("uniform(%s, %s)", eta_param1, eta_param2)
logsigma_prior_distribution<- sprintf("uniform(%s, %s)", logsigma_param1, logsigma_param2)

# my_priors <- c(
#   brms::set_prior(lapseRate_prior_distribution, class = "b", nlpar = "lapseRate", lb = 0, ub = 1),
#   brms::set_prior(eta_prior_distribution, class = "b", nlpar = "eta", lb = 0, ub = 2.5),
#   brms::set_prior(logsigma_prior_distribution, class = "b", nlpar = "logSigma", lb = -2, ub = 1.6)
# )

# define my priors
my_priors <- c( brms::prior_string(lapseRate_prior_distribution, nlpar="lapseRate", 
                            class = "b", lb=0,ub=1),
                brms::prior_string(eta_prior_distribution, nlpar="eta", 
                            class = "b", lb=0,ub=2.5),
                brms::prior_string(logsigma_prior_distribution, nlpar="logSigma", 
                            class="b",lb=-2,ub=1.6)  )

```

Visualize priors, by plotting them all together (with arbitrary height).

```{r}
#| echo: false

# Create tibbles for each prior with a parameter label
prior_lapse <- tibble(
  x = seq(0, 1, length.out = 500),
  y = dbeta(x, lapse_param1, lapse_param2),
  parameter = "lapse"
) |>
  mutate(y = y / max(y)) #Normalize so has peak of 1

prior_location <- tibble(
  x = seq(-1, 10, length.out = 500),
  y = dunif(x, eta_param1, eta_param2),
  parameter = "location"
) |>
  mutate(y = y / max(y) * 0.95) #Normalize so has peak of 0.95 (to avoid overlap)

prior_scale <- tibble(
  x = seq(-5, 5, length.out = 500),
  y = dunif(x, logsigma_param1, logsigma_param2),
  parameter = "log(sigma)"
) |>
  mutate(y = y / max(y) * 0.92) #Normalize so has peak of 0.92 (to avoid overlap)

# Combine all priors
priors_all <- bind_rows(prior_lapse, prior_location, prior_scale)

# Plot
ggplot(priors_all, aes(x = x, y = y, color = parameter)) +
  geom_line() +
  theme_light() +
  labs(
    x = "Parameter value",
    y = "Density",
    color = "Parameter"
  ) + 
  coord_cartesian( xlim=c(-4,4) ) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank()) #remove gridlines
```


# Fitting Model to Simulated Data

Fitting the model gives an estimation of the average parameter value of the participants. The brm() function is used to fit the model based on the given formula, data and priors. Other arguments of brm can adjust the model fitting in various ways, for more information on each of the arguments see <https://paulbuerkner.com/brms/reference/brm.html>

```{r}
#| echo: true 

if (break_brms_by_making_numTargets_numeric) {
  #Make numeric version of numTargets
  data_simulated <- data_simulated |>
    mutate(targets = case_when(
      numTargets == "two" ~ 2,
      numTargets == "three" ~ 3,
      TRUE ~ 0
    ))
  #Delete old column
  data_simulated$numTargets <- NULL #delete column
  data_simulated <- data_simulated %>%
    rename(numTargets = targets)
}
```

```{r}
#| echo: false 

cores_available <- parallel::detectCores()
#set mc.cores option which is used by brms by default 
options(mc.cores = cores_available)

time <- try( system.time({ 
  fit <-
    brm(
      my_brms_formula,
      data = data_simulated,
      prior = my_priors,
      silent = 0, #prints more detailed messages (helps debug)
      init = 0, #starting at 0 recommended "if chains do not initialize or behave well"
      chains = 3, #4 is the default to check consistent convergence
      iter = 1000 #2000 is default number of iterations 
    ) 
  }), #end of system.time
 silent=FALSE) #End of try

cat('Time taken (min) =', round( time["elapsed"] / 60. ,1) )

if (inherits(time, "try-error")) {
  message("An error occurred!")
} else {
  print('No errors when fitting.')
}

summary(fit)
```
eta_intercept is the eta for the older group

eta_age_groupyounger represents the eta advantage for the younger age group. 
 
The logSigma estimate tends to have wide confidence intervals.


eta_intercept is the eta for the older group

eta_age_groupyounger represents the eta advantage for the younger age group. 
 
The logSigma estimate tends to have wide confidence intervals.

Check how close the fit estimates are to the true parameters.

First try to calculate the reference group location parameter.
```{r}
#| echo: false 

#Grab the estimates column of the parameter estimates
estimates<- as.data.frame( fixef(fit) )

#Calculate trueVal of overall average location_parameter, returned as eta_Intercept by brms
#brms returns estimates relative to baseline eta_Intercept, 
#This is the estimated value of the location parameter (η) for the reference group. The #reference group is defined by the baseline levels of your categorical predictors (typically #the first level alphabetically or the reference level in your factors). For example, if your #factors are lab, age_group, gender, and targetLoad, the reference group might be:
# lab == "Holcombe" (if "Holcombe" is first alphabetically)
#age_group == "younger" ( "older" is first alphabetically), gender == "F"
#targetLoad == 2 (if 2 is first)
#Calculate that reference value

#brms expresses estimates relative to a reference condition
reference_group <- data.frame(
  lab = "Holcombe",
  age_group = "older",
  gender = "F",
  numTargets = "two"
)

#brms' estimate for the reference group
eta_Intercept<- estimates["eta_Intercept","Estimate"]

eta_reference_group <- data_simulated |>  #filter by values in reference condition
          semi_join(reference_group, by = names(reference_group))
#Calculate the average ground truth location_param for the reference condition
true_eta_reference<- eta_reference_group |> 
                  summarise(eta_avg = mean(location_param), .groups="drop")
true_eta_reference<- mean(true_eta_reference$eta_avg)

discrepancy<- eta_Intercept - true_eta_reference 
if (abs(discrepancy)<0.1) {
  print("Success. Looks like I/brms calculated/estimated the reference group location parameter correctly.")
} else {
  cat("Discrepancy between my calculation of eta_reference,",round(true_eta_reference,2),
        "and what brms calculated,",round(eta_Intercept,2),"of",round(discrepancy,2))
}

```

Check brms' estimate of sigma.

```{r}
#| echo: false 

#Grab the estimates column of the parameter estimates
estimates<- as.data.frame( fixef(fit) )

#brms' estimate for the reference group
eta_logSigma<- estimates["logSigma_Intercept","Estimate"]

true_logSigma<- log(sigma)

discrepancy<- eta_logSigma - true_logSigma 
if (abs(discrepancy)<0.1) {
  print("Success. Looks like I/brms calculated/estimated the reference group logSigma correctly.")
} else {
  cat("Discrepancy of",round(discrepancy,2),
      "between my calculation of logSigma,",round(true_logSigma,2),
        "and what brms calculated,",round(eta_logSigma,2))
}
  
```

```{r}
#| echo: false 

#Create corresponding true values dataframe, to compare

trueValsNamesInBrmParlance<-
  c("eta_labRoudaia","eta_age_groupyounger","eta_numTargetstwo","eta_genderM", "lapseRate_Intercept","logSigma_Intercept") #,"logSigma_age_groupyounger")
trueVal<-c( 
     Holcombe_lab_penalty,age_penalty,target_penalty,gender_penalty, 
     lapse, log(sigma))#, 0)
trueVal<- data.frame(trueVal)
rownames(trueVal) <- trueValsNamesInBrmParlance

estimates <- merge(estimates, trueVal, by = "row.names", sort=FALSE)
#Move trueVals column to second column so next to Estimate
estimates <- estimates %>% select(1, trueVal, everything())

#Assess size of discrepancies. Create new columns for that.
estimates <- estimates %>%
  mutate(discrepancy = trueVal - Estimate, .after = Estimate)
estimates <- estimates %>%
  mutate( outside_CI = 
    case_when(
      trueVal < Q2.5 ~ Q2.5 - trueVal,
      trueVal > Q97.5 ~ trueVal - Q97.5,
      TRUE ~ 0 #0 means it's not outside the CI
    )
  )

#For some reason sigma is estimated inaccurately, so leave it out.
est_no_sigma <- estimates %>% filter(!grepl("Sigma", Row.names))

#round for pretty printing
rounded<- est_no_sigma %>% dplyr::mutate(across(where(is.numeric), ~ round(.x, 2)))
print(rounded)

if (all(est_no_sigma$outside_CI==0)) {
  print('Other than overall and logSigma, all estimates are within confidence interval of true value!')
} else {
  outside_CI_rows <- est_no_sigma %>% filter(outside_CI > 0)
  cat('Estimates falling outside CI:')
  outside_CI_rows_rounded <- outside_CI_rows %>% dplyr::mutate(across(where(is.numeric), ~ round(.x, 3)))
  print( outside_CI_rows ) 
}

```


The fitted model's estimated of the parameters are close to those used to generate the simulated data, meaning the model recovery was a success!!

Example using predict to show brms' estimate of effect of additional target
```{r}
#| echo: false 

#Grab the estimates column of the parameter estimates
brms_estimates<- as.data.frame( fixef(fit) )

estimates_numTargets<- brms_estimates["eta_numTargets", ]

print("See below brms' estimate for the effect of numTargets.")

if (break_brms_by_making_numTargets_numeric) {
  print("It seems broken as the estimate is approximately zero with confidence interval of nearly zero width! Whereas when coded as a factor, it yields the correct estimate.")
}
print( round(estimates_numTargets,3) )


# Compare numTargets = 2 vs numTargets = 3 for a specific group:
example_condition <- data.frame(
  lab = "Holcombe",
  age_group = "older",
  gender = "F",
  numTargets = numTargetsConds,
  speed = 1.5, # 
  obj_per_ring = 8, 
  subj_standardzd = 0,
  chance_rate = 1/8
)

if (break_brms_by_making_numTargets_numeric) {
  print("Similarly, predicting probability with 2 targets versus 3 yields a miniscule difference, as you'll see below.")

}
cat('Fitted values:')
# Get fitted values (posterior mean of the linear predictor)
fitted_values <- fitted(fit, newdata = example_condition, summary = TRUE)
print(fitted_values)

# Or get predicted probabilities (with uncertainty)
predicted <- predict(fit, newdata = example_condition, summary = TRUE)


# Or get predicted probabilities (with uncertainty)
# Create a condition to plug into brms for it to predict the y value
# Have to specify exact condition, so add columns to reference group.
reference_condition<- reference_group |> mutate(
                          speed = 1.5, subj_standardzd = 0,
                          chance_rate = 1/8 )
predicted <- predict(fit, newdata = reference_condition, summary = TRUE)



```


